---
title: "MUSA508_midterm"
author: "Jasmine Siyu Wu"
date: "10/18/2021"
output: 
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---
<br />

```{r knitr.setup, include=TRUE, cache = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 10, fig.height = 5)
```
<br />

## 1. Introduction
[placeholder of paragraphs]

<br />
<br />


## 2. Setting Up Environment and Styling
This section sets up the environment including libraries, styling options, quintile break functions, and color palettes.

**Loading Libraries**
```{r setup, include=TRUE, cache = FALSE, message = FALSE}
library(tidyverse)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot)
library(kableExtra)
library(jtools)     # for regression model plots
library(ggstance)
library(osmdata)
library(knitr)
library(tidycensus)
library(scales)
library(stargazer)


options(scipen=999) 
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)

# functions and data directory
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")
```

**Loading Styling Options**
```{r load_styling, include=TRUE, cache = FALSE, message = FALSE, warning = FALSE}
mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 16,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.text.x = element_text(size = 14))
}

plotTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 16,colour = "black"),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2),
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=12),
    axis.title = element_text(size=12),
    axis.text = element_text(size=10),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "italic"),
    legend.text = element_text(colour = "black", face = "italic"),
    strip.text.x = element_text(size = 14)
  )
}
```

**Loading Quantile Break Functions**
```{r qBr, include=TRUE, cache = FALSE, message = FALSE, warning = FALSE}
qBr <- function(df, variable, rnd) {
 if (missing(rnd)) {
    as.character(quantile(round(df[[variable]],0),
                 c(.01,.2,.4,.6,.8), na.rm=T))
 } else if (rnd == FALSE | rnd == F) {
    as.character(formatC(quantile(df[[variable]],
                 c(.01,.2,.4,.6,.8), na.rm=T), digits = 3))
 }
}


q5 <- function(variable) {as.factor(ntile(variable, 5))}
```

**Loading Hexadecimal Color Palette**
```{r palette5, include=TRUE, cache = FALSE, message = FALSE, warning = FALSE}
palette5 <- c("#eff3ff", "#bdd7e7", "#6baed6", "#3182bd", "#08519c")
```

<br />
<br />

## 3. Data Wrangling
[placeholder of paragraphs]

<br />

### 3.1. Loading Data

loading features
```{r read_data, cache = TRUE, message = FALSE, warning = FALSE}
# read in boulder training data
boulder.sf <- 
  st_read("studentData.geojson", crs="ESRI:102254") %>%
  st_transform('ESRI:102254') %>%  # "NAD_1983_HARN_StatePlane_Colorado_North_FIPS_0501"
  st_make_valid() %>% 
  mutate(Age = 2021 - builtYear,
         PricePerSF = price/TotalFinishedSF)

boulder.sf <- transform(boulder.sf, qualityCode = as.character(qualityCode), 
                        bldgClass = as.character(bldgClass),
                        ConstCode = as.character(ConstCode),
                        Ac = as.character(Ac),
                        Heating = as.character(Heating),
                        ExtWallPrim = as.character(ExtWallPrim),
                        ExtWallSec = as.character(ExtWallSec),
                        IntWall = as.character(IntWall),
                        Roof_Cover = as.character(Roof_Cover))


boulder.sf[2638,]$price <- 315000 #mistake input as 31500000, checked on zillow



# read in commissioner districts
# data source: boulder county open data portal
coms_dists.sf <-
  st_read("https://opendata.arcgis.com/datasets/1fb6698cad044e3fb0471d1d464049ad_0.geojson") %>%
  st_transform(st_crs(boulder.sf))

coms_dists.sf <- transform(coms_dists.sf, DISTRICT = as.character(DISTRICT))

# read in Boulder County municipities
# data source: boulder county open data portal
admin.sf <-
  st_read("https://opendata.arcgis.com/datasets/9597d3916aba47e887ca563d5ac15938_0.geojson") %>%
  st_transform(st_crs(boulder.sf)) %>% dplyr::select(ZONECLASS, ZONEDESC)

#admin.unin <- st_combine(st_difference(county_boundary, admin.sf)) %>% mutate(ZONECLASS = "UNIN", ZONEDESC = "Unincorporated")  %>%  rename(geometry=x)


# read in boulder county playgrounds (point)
# data source: boulder county open data portal
playgrounds.sf <- 
  st_read("https://opendata.arcgis.com/datasets/b89ea27bc3cd492682503f03df1a9fb9_0.geojson") %>%
  st_as_sf(coords=playgrounds.sf$geometry, crs = 4326) %>% 
  st_transform(st_crs(boulder.sf)) 


# read in schools in boulder county from open street map
# data source: open street map
schools <- osmdata_sf(getbb("Boulder County", base_url="https://nominatim.openstreetmap.org") %>% 
                             opq() %>% 
                             add_osm_feature("amenity", "school"))$osm_points

schools.sf <- st_as_sf(schools, coords=schools$geometry, crs = 4326) %>% 
  st_geometry(schools.sf$geometry) %>%
  st_transform(st_crs(boulder.sf))  %>%
  st_sf() %>%
  cbind(., schools$name) %>%
  rename(NAME = schools.name)


# read in parks from open street map
# data source: open street map
parks <- osmdata_sf(getbb("Boulder County", base_url="https://nominatim.openstreetmap.org") %>% opq() %>%
                      add_osm_feature(key = 'leisure', value = "park"))

parks.sf <- st_geometry(parks$osm_polygons) %>%
  st_transform(st_crs(boulder.sf)) %>%
  st_sf() %>%
  cbind(., parks$osm_polygons$name) %>%
  rename(NAME = parks.osm_polygons.name)


# read in pollution score


# read in transit stops
# data source: open street map
bus_stops <- osmdata_sf(getbb("Boulder County", base_url="https://nominatim.openstreetmap.org") %>% 
                             opq() %>% 
                             add_osm_feature("highway", "bus_stop"))$osm_points

bus_stops.sf <- st_as_sf(bus_stops, coords=bus_stops$geometry, crs = 4326) %>% 
  st_geometry(bus_stops.sf$geometry) %>%
  st_transform(st_crs(boulder.sf)) %>%
  st_sf() %>%
  cbind(., bus_stops$name) %>%
  rename(NAME = bus_stops.name)


# read in highway ramps



# read in grass land from open street map
# data source: open street map
grass <- osmdata_sf(getbb("Boulder County", base_url="https://nominatim.openstreetmap.org") %>% opq() %>%
                      add_osm_feature(key = 'landuse', value = "grass"))

grass.sf <- st_geometry(grass$osm_polygons) %>%
  st_transform(st_crs(boulder.sf)) %>%
  st_sf() %>%
  cbind(., grass$osm_polygons$name) %>%
  rename(NAME = grass.osm_polygons.name)



# read in lakes and reservoirs data
# data source: boulder county open data portal
lakes.sf <-
  st_read("https://opendata.arcgis.com/datasets/e0759cd6dc8f4990a04b77a29c988b55_0.geojson") %>%
  st_transform(st_crs(boulder.sf))


# read in trail routes
# data source: boulder county open data portal
trails.sf <-
  st_read("https://opendata.arcgis.com/datasets/3ed1045255c34eada007840a5dd84de9_0.geojson") %>%
  st_transform(st_crs(boulder.sf))


# read in trailheads
# data source: boulder county open data portal
trailheads.sf <-
  st_read("https://opendata.arcgis.com/datasets/3a950053bbef46c6a3c2abe3aceee3de_0.geojson") %>%
  st_transform(st_crs(boulder.sf))


# read in parcels
# data source: boulder county open data portal
#parcels.sf <-
#  st_read("https://opendata.arcgis.com/datasets/89ae49d4ddf246388ee5f5e952aa84db_0.geojson") %>%
#  st_transform(st_crs(boulder.sf))


# read in mobile homes
# data source: boulder county open data portal
trailheads.sf <-
  st_read("https://opendata.arcgis.com/datasets/8d53540bcc264bc3b13a500ab274bfa6_0.geojson") %>%
  st_transform(st_crs(boulder.sf))

```
<br />

Loading boundaries and demographic data

```{r read_acs_data, cache = TRUE, message = FALSE, warning = FALSE}
acs_variable_list.2019 <- load_variables(2019, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)



# read in boulder county boundary
# data source: boulder county open data portal
county_boundary <-
  st_read("https://opendata.arcgis.com/datasets/964b8f3b3dbe401bb28d49ac93d29dc4_0.geojson") %>%
  st_transform(st_crs(boulder.sf))

# read in places (cities)
# data source: ACS 2015-19 5yr
places <- get_acs(geography = "place", 
                    year = 2019,
                    state = 08,
                    variables = (TotalPop = 'B01001_001E'),
                    survey = "acs5",
                    output = "wide",
                    geometry = TRUE) %>%
  st_transform(st_crs(boulder.sf))

# filter out cities in boulder county
bc.places <-  
  st_centroid(places)[county_boundary,] %>%
  st_drop_geometry() %>%
  left_join(places) %>%
  st_sf() 

# read in ACS 5yr 2015-19 estimates
# variables: population density, poverty rate, median household income, white population percent, 
tractsBC19 <- 
  get_acs(geography = "tract", variables = c("B01001_001","B02001_002","B15001_050",
                                             "B15001_009","B19013_001","B25058_001",
                                             "B06012_002"), 
          year=2019, state=08, county=013, geometry=T,output="wide") %>%
  st_transform(st_crs(boulder.sf)) %>%
  rename(TotalPop = B01001_001E, 
         Whites = B02001_002E,
         FemaleBachelors = B15001_050E, 
         MaleBachelors = B15001_009E,
         MedHHInc = B19013_001E, 
         MedRent = B25058_001E,
         TotalPoverty = B06012_002E) %>%
  dplyr::select(-NAME, -starts_with("B")) %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop,0), # avoid of NAs
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop),0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0), 
         areaSqMi = as.numeric(st_area(.))*0.00000038610215855, 
         popDensity = TotalPop/areaSqMi,
         raceContext = ifelse(pctWhite > .5, "Majority White", "Majority Non-White"),
         incomeContext = ifelse(MedHHInc > 83019, "High Income", "Low Income"),
         year = "2019") %>%
  dplyr::select(-Whites, -FemaleBachelors, -MaleBachelors, -TotalPoverty) 




# visualize raw data
ggplot() +
    geom_sf(data=county_boundary, show.legend = NA, color='grey',alpha=0.5, lwd=0.3)+
    geom_sf(data=bc.places, show.legend = NA, color='grey', lwd=0.1)+
    geom_sf(data = boulder.sf, aes(colour = q5(price)), 
            show.legend = "point", size = 1) + 
    scale_fill_manual(values = palette5,
                      labels = qBr(boulder.sf, "price"),
                      name = "House Price\n(Quintile Breaks)")
    #geom_sf(data=playgrounds.sf, color="black", alpha=0.7, shape=18, size=1)+
    #geom_sf(data=schools, color="black", shape=15, alpha=0.7, size=1) 
    #+    coord_sf(datum=st_crs(boulder.sf))

```

<br />

### 3.2. Feature Engineering

```{r feature_engineering, cache = TRUE, message = FALSE, warning = FALSE}
# number of features within 0.5 mile

boulder.sf <- 
  mutate(boulder.sf, 
         playground.Buffer=lengths(st_intersects(st_buffer(boulder.sf, 1609/2), playgrounds.sf))) %>% 
  mutate(boulder.sf, 
         school.Buffer=lengths(st_intersects(st_buffer(boulder.sf, 1609/2), schools.sf))) %>%
  mutate(boulder.sf, 
         parks.Buffer=lengths(st_intersects(st_buffer(boulder.sf, 1609/2), parks.sf)))


# average nearest neighbor distance to 1-NN and 3-NN parks
st_c <- st_coordinates #st_c(coordinates)

boulder.sf <-
  boulder.sf %>% 
    mutate(park_nn1 = nn_function(st_c(boulder.sf), st_c(st_centroid(parks.sf)), 1)/1609,
           park_nn3 = nn_function(st_c(boulder.sf), st_c(st_centroid(parks.sf)), 3)/1609) #in mile

#mean(boulder.sf$park_nn1)
#mean(boulder.sf$park_nn3)


# average nearest neighbor distance to 1-NN and 3-NN schools
boulder.sf <-
  boulder.sf %>% 
    mutate(school_nn1 = nn_function(st_c(boulder.sf), st_c(schools.sf), 1)/1609, 
           school_nn3 = nn_function(st_c(boulder.sf), st_c(schools.sf), 3)/1609) # in mile

#mean(boulder.sf$school_nn1)
#mean(boulder.sf$school_nn3)


# average nearest neighbor distance to 1-NN and 3-NN playgrounds
boulder.sf <-
  boulder.sf %>% 
    mutate(playground_nn1 = nn_function(st_c(boulder.sf), st_c(playgrounds.sf), 1)/1609, 
           playground_nn3 = nn_function(st_c(boulder.sf), st_c(playgrounds.sf), 3)/1609) # in mile

#mean(boulder.sf$playground_nn1)
#mean(boulder.sf$playground_nn3)


# average nearest neighbor distance to 1-NN and 3-NN lakes
boulder.sf <-
  boulder.sf %>% 
    mutate(lake_nn1 = nn_function(st_c(boulder.sf), st_c(st_centroid(lakes.sf)), 1)/1609, 
           lake_nn3 = nn_function(st_c(boulder.sf), st_c(st_centroid(lakes.sf)), 3)/1609) # in mile

#mean(boulder.sf$lake_nn1)
#mean(boulder.sf$lake_nn3)


# average nearest neighbor distance to 1-NN and 3-NN parks trail heads
#boulder.sf <-
#  boulder.sf %>% 
#    mutate(trailhd_nn1 = nn_function(st_c(boulder.sf), st_c(trailheads.sf), 1)/1609,
#           trailhd_nn3 = nn_function(st_c(boulder.sf), st_c(trailheads.sf), 3)/1609) #in mile

#mean(boulder.sf$trailhd_nn1)
#mean(boulder.sf$trailhd_nn3)


# attach demographic and socioeconomic data to house points
boulder.sf <- 
  st_join(boulder.sf,
          dplyr::select(tractsBC19, GEOID, MedRent, TotalPop, pctWhite, pctBachelors, pctPoverty, popDensity, raceContext, incomeContext))

boulder.sf <- 
  st_join(boulder.sf,
          dplyr::select(admin.sf, ZONECLASS, ZONEDESC)) 
# %>% mutate(ZONEDESC = replace_na(ZONEDESC, "Unincorporated"),
#         ZONECLASS = replace_na(ZONECLASS, "UNIN"))

boulder.sf <- 
  st_join(boulder.sf,
          dplyr::select(coms_dists.sf, DISTRICT, Representative)) 

```

<br />

### 3.3. Exploratory Data Analysis

```{r price_map, fig.height=5, fig.width=10, cache = TRUE, message = FALSE, warning = FALSE}
# ggplot, reorder

# Mapping data
ggplot() +
  geom_sf(data=county_boundary, show.legend = NA, color='grey',alpha=0.5, lwd=0.3)+
  geom_sf(data=admin.sf, show.legend = NA, color='grey', lwd=0.1)+
  geom_sf(data = boulder.sf, aes(colour = q5(PricePerSF)), 
          show.legend = "point", size = 0.8) +
  scale_colour_manual(values = palette5,
                   labels=qBr(boulder.sf,"PricePerSF"),
                   name="Quintile\nBreaks") +
  labs(title="Price Per Square Foot, Boulder County, CO") +
  mapTheme()

# method to build html slider
```


<br />
<br />


## 4. Correlation Analysis
[placeholder of paragraphs]

<br />

### 4.1. Analyzing Associations


```{r Corellation, cache = TRUE, message = FALSE, warning = FALSE，fig.height = 4, fig.width=12, }

## Home Features cor
st_drop_geometry(boulder.sf) %>% 
  dplyr::select(price, TotalFinishedSF, Age, mainfloorSF) %>%
  #filter(SalePrice <= 1000000, Age < 500) %>% 
  gather(Variable, Value, -price) %>%  
   ggplot(aes(Value, price)) + # the x and y axes are 3 variables and sale price
     geom_point(size = .5) + 
     geom_smooth(method = "lm", se=F, colour = "#FA7800") + # geom_smooth is the best fit line (method is linear regression)
     facet_wrap(~Variable, ncol = 3, scales = "free") +
     labs(title = "Price as a function of continuous variables") +
     plotTheme()
```


```{r crime_corr, fig.height = 4, fig.width=12, cache = TRUE, message = FALSE, warning = FALSE}
## park cor
boulder.sf %>%
  st_drop_geometry() %>%
  dplyr::select(price, starts_with("park_")) %>%
  gather(Variable, Value, -price) %>% 
   ggplot(aes(Value, price)) +
     geom_point(size = .5) + geom_smooth(method = "lm", se=F, colour = "#FA7800") +
     facet_wrap(~Variable, nrow = 1, scales = "free") +
     labs(title = "Price as a function of continuous variables") +
     plotTheme()
```

<br />

### 4.2. Correlation Matrix


```{r correlation_matrix, fig.height=10, fig.width=10, cache = TRUE, message = FALSE, warning = FALSE}
numericVars <- 
  select_if(st_drop_geometry(boulder.sf), is.numeric) %>% 
  dplyr::select(-MUSA_ID, -toPredict, -builtYear, -Stories, -UnitCount) %>% 
  na.omit() 

# only for continuous variables
ggcorrplot(  #correlation plot
  round(cor(numericVars), 1), 
  p.mat = cor_pmat(numericVars),
  show.diag = TRUE,
  colors = c("#25CB10", "white", "#FA7800"),
  type="lower",
  insig = "blank") +  
    labs(title = "Correlation across numeric variables") 

# noting strongly associates with sales price
# house size variables are closely related --- keep one or engineer them
# crime variables are also closely related
```


<br />
<br />


## 5. Regression Analysis
[placeholder of paragraphs]


<br />

### 5.1. Univariate Regression 1

```{r simple_reg, cache = TRUE, message = FALSE, warning = FALSE}
areaReg <- lm(price ~ TotalFinishedSF, data = boulder.sf)

summary(areaReg)
```
**Prediction example**

```{r calculate prediction, cache = TRUE, message = FALSE, warning = FALSE}
new_TotalFinishedSF = 4000

# "by hand"
152932.565 + 303.609 * new_TotalFinishedSF

# predict() function
predict(areaReg, newdata = data.frame(TotalFinishedSF = 4000))
```

**Plot of marginal regression response**
```{r effectPlot, cache = TRUE, message = FALSE, warning = FALSE}
effect_plot(areaReg, pred = TotalFinishedSF, interval = TRUE, plot.points = TRUE)
```


<br />

### 5.2. Multivariate Regression

```{r mutlivariate_regression, cache = TRUE, message = FALSE, warning = FALSE}
# can play with categorical variables?
reg1 <- lm(price ~ ., data = st_drop_geometry(boulder.sf) %>% 
                             dplyr::select(price, Age, qualityCode, ConstCode, bsmtType, #designCode, 
                                    carStorageSF, nbrBedRoom, TotalFinishedSF, mainfloorSF, Roof_Cover, #nbrRoomsNobath,
                                    nbrFullBaths, parks.Buffer, #playground.Buffer, school.Buffer,
                                    school_nn3, playground_nn1, lake_nn3, #park_nn3, 
                                    popDensity, pctBachelors))
           
summary(reg1)
```

**Cross Validation**
```{r cross_validation, cache = TRUE, message = FALSE, warning = FALSE}
fitControl <- trainControl(method = "cv", number = 100)
set.seed(825)

boulder_toPdt <- subset(boulder.sf, boulder.sf$toPredict==1)
boulder_pdt <- subset(boulder.sf, boulder.sf$toPredict==0)

reg1.cv <- 
  train(price ~ ., data = st_drop_geometry(boulder_pdt) %>% 
                                dplyr::select(price,Age, qualityCode, ConstCode, bsmtType, #designCode, 
                                    carStorageSF, nbrBedRoom, TotalFinishedSF, mainfloorSF, Roof_Cover, #nbrRoomsNobath,
                                    nbrFullBaths, parks.Buffer, #playground.Buffer, school.Buffer,
                                    school_nn3, playground_nn1, lake_nn3, #park_nn3, 
                                    popDensity, pctBachelors), 
     method = "lm", trControl = fitControl, na.action = na.pass)

reg1.cv

```


**Distribution of MAE**
```{r hist_MAE, cache = TRUE, message = FALSE, warning = FALSE}
reg1.cv$resample$MAE

ggplot(data=reg1.cv$resample, aes(reg1.cv$resample$MAE)) + 
  geom_histogram(xlab = "Mean Absolute Error", ylab = "Count",  
                 color="orange", fill="orange", bins=50) +
  labs(title = "Distribution of MAE",
       subtitle = "k-fold cross validation; k=100") +
  plotTheme()

# subtitle = "k-fold cross validation; k=100",
```


### 5.3. Baseline Regression

```{r baseline_regression, cache = TRUE, message = FALSE, warning = FALSE}
inTrain <- createDataPartition(
  y = paste(boulder_pdt$Roof_Cover,
            boulder_pdt$qualityCode, boulder_pdt$Ac), 
  p = .75, list = FALSE) %>%
  append(which(boulder_pdt$qualityCode == "10")) %>%
  append(which(boulder_pdt$GEOID == "08013012300")) %>%
  append(which(boulder_pdt$ConstCode %in% c("410", "420"))) %>%
  append(which(boulder_pdt$bsmtType == "LWU"))

boulder.training <- boulder_pdt[inTrain,]
boulder.test <- boulder_pdt[-inTrain,]
#boulder.training <- subset(boulder.sf, boulder.sf$toPredict==0) 
#boulder.test <- subset(boulder.sf, boulder.sf$toPredict==1)

reg.training <- 
  lm(price ~ ., data = as.data.frame(boulder.training) %>% 
       dplyr::select(price, Age, qualityCode, ConstCode, bsmtType, #designCode, 
                     carStorageSF, nbrBedRoom, TotalFinishedSF, mainfloorSF, #nbrRoomsNobath,
                     nbrFullBaths, parks.Buffer, #playground.Buffer, school.Buffer,
                     school_nn3, playground_nn1, lake_nn3, #park_nn3, 
                     popDensity, pctBachelors))
summary(reg.training)


#boulder.training <-
#  boulder.training %>%
#  mutate(Regression = "Baseline Regression",
#         price.Predict = predict(reg.training, boulder.training),
#         price.Error = price.Predict - price,
#         price.AbsError = abs(price.Predict - price),
#         price.APE = (abs(price.Predict - price))/price.Predict)


boulder.test <-
  boulder.test %>%
  mutate(Regression = "Baseline Regression",
         price.Predict = predict(reg.training, boulder.test),
         price.Error = price.Predict - price,
         price.AbsError = abs(price.Predict - price),
         price.APE = (abs(price.Predict - price))/price.Predict)

```


<br />
<br />


## 6. Neighborhood Model
[placeholder of paragraphs]

### 6.1. Spatial Lag

```{r spatial_lag, cache = TRUE, message = FALSE, warning = FALSE}

coords <- st_coordinates(boulder.sf) 
neighborList <- knn2nb(knearneigh(coords, 5))
spatialWeights <- nb2listw(neighborList, style="W")
boulder.sf <- mutate(boulder.sf, lagPrice=lag.listw(spatialWeights, boulder.sf$price, zero.policy=TRUE, NAOK = TRUE)) 


#coords.training <-  st_coordinates(boulder.training)
#neighborList.training <- knn2nb(knearneigh(coords.training, 5))
#spatialWeights.training <- nb2listw(neighborList.training, style="W")
#boulder.training %>% 
#  mutate(lagPriceError = lag.listw(spatialWeights.training, price.Error, zero.policy=TRUE, NAOK = TRUE)) %>%
#  ggplot(aes(lagPriceError, price.Error))


coords.test <-  st_coordinates(boulder.test) 
neighborList.test <- knn2nb(knearneigh(coords.test, 5))
spatialWeights.test <- nb2listw(neighborList.test, style="W")
boulder.test <- 
  mutate(boulder.test, 
         lagPrice = lag.listw(spatialWeights.test, boulder.test$price, zero.policy=TRUE, NAOK = TRUE),
         lagPriceError = lag.listw(spatialWeights.test, boulder.test$price.Error, zero.policy=TRUE, NAOK = TRUE))
```


### 6.2. Moran's I

observed Moran's I > 0, so clustering

```{r moransI, cache = TRUE, message = FALSE, warning = FALSE}
moranTest <- moran.mc(boulder.test$price.Error, 
                      spatialWeights.test, nsim = 999)

ggplot(as.data.frame(moranTest$res[c(1:999)]), aes(moranTest$res[c(1:999)])) +
  geom_histogram(binwidth = 0.01) +
  geom_vline(aes(xintercept = moranTest$statistic), colour = "#FA7800",size=1) +
  scale_x_continuous(limits = c(-1, 1)) +
  labs(title="Observed and permuted Moran's I",
       subtitle= "Observed Moran's I in orange",
       x="Moran's I",
       y="Count") +
  plotTheme()
```


### 6.3. Neighborhood Model

```{r neighborhood_model, cache = TRUE, message = FALSE, warning = FALSE}
# no neighbohood so that usually using census tract
# https://www.bouldercounty.org/families/disease/covid-19-information/covid-19-vaccine-information/vaccine-data/
left_join(
  st_drop_geometry(boulder.test) %>%
    group_by(DISTRICT) %>%
    summarize(meanPrice = mean(price, na.rm = T)),
  mutate(boulder.test, predict.fe = predict(lm(price ~ DISTRICT, data = boulder.test), boulder.test)) %>%
    st_drop_geometry %>%
    group_by(DISTRICT) %>%
      summarize(meanPrediction = mean(predict.fe))) %>%
      kable() %>% kable_styling()

reg.nhood <- lm(price ~ ., data = as.data.frame(boulder.training) %>% 
                                  dplyr::select(DISTRICT, price, Age, qualityCode, ConstCode, bsmtType, #designCode, 
                                                carStorageSF, nbrBedRoom, TotalFinishedSF, mainfloorSF, #nbrRoomsNobath,
                                                nbrFullBaths, parks.Buffer, #playground.Buffer, school.Buffer,
                                                school_nn3, playground_nn1, lake_nn3, #park_nn3, 
                                                popDensity, pctBachelors))
summary(reg.nhood)

boulder.test.nhood <-
  boulder.test %>%
  mutate(Regression = "Neighborhood Effects",
         price.Predict = predict(reg.nhood, boulder.test),
         price.Error = price.Predict- price,
         price.AbsError = abs(price.Predict- price),
         price.APE = (abs(price.Predict- price)) / price)
```


**Accuracy of the neighborhood model**

```{r accuracy_statistics, cache = TRUE, message = FALSE, warning = FALSE}
bothRegressions <- 
  rbind(
    dplyr::select(boulder.test, starts_with("price"), Regression, DISTRICT) %>%
      mutate(lagPriceError = lag.listw(spatialWeights.test, price.Error)),
    dplyr::select(boulder.test.nhood, starts_with("price"), Regression, DISTRICT) %>%
      mutate(lagPriceError = lag.listw(spatialWeights.test, price.Error)))   

st_drop_geometry(bothRegressions) %>%
  gather(Variable, Value, -Regression, -DISTRICT) %>%
  filter(Variable == "price.AbsError" | Variable == "price.APE") %>%
  group_by(Regression, Variable) %>%
    summarize(meanValue = mean(Value, na.rm = T)) %>%
    spread(Variable, meanValue) %>%
    kable()
```



```{r accuracy_plot, cache = TRUE, message = FALSE, warning = FALSE, fig.height=4, fig.width=8}
bothRegressions %>%
  dplyr::select(price.Predict, price, Regression) %>%
    ggplot(aes(price, price.Predict)) +
  geom_point() +
  stat_smooth(aes(price, price), 
             method = "lm", se = FALSE, size = 1, colour="#FA7800") + 
  stat_smooth(aes(price.Predict, price), 
              method = "lm", se = FALSE, size = 1, colour="#25CB10") +
  facet_wrap(~Regression) +
  labs(title="Predicted sale price as a function of observed price",
       subtitle="Orange line represents a perfect prediction; Green line represents prediction") +
  plotTheme()
```



**Generalizability of the neighborhood model**

```{r mape_plot, cache = TRUE, message = FALSE, warning = FALSE, fig.height=4, fig.width=8}
st_drop_geometry(bothRegressions) %>%
  group_by(Regression, DISTRICT) %>%
  summarize(mean.MAPE = mean(price.APE, na.rm = T)) %>%
  ungroup() %>% 
  left_join(coms_dists.sf) %>%
    st_sf() %>%
    ggplot() + 
      geom_sf(aes(fill = mean.MAPE)) +
      geom_sf(data = bothRegressions, colour = "black", size = .5) +
      facet_wrap(~Regression) +
      scale_fill_gradient(low = palette5[1], high = palette5[5],
                          name = "MAPE") +
      labs(title = "Mean test set MAPE by neighborhood") +
      mapTheme()
```

### 6.4. Context Neighborhood Model


```{r context_map, cache = TRUE, message = FALSE, warning = FALSE, fig.height=4, fig.width=8}
grid.arrange(ncol = 2,
  ggplot() + geom_sf(data = na.omit(tractsBC19), aes(fill = raceContext)) +
    scale_fill_manual(values = c("#25CB10", "#FA7800"), name="Race Context") +
    labs(title = "Race Context") +
    mapTheme() + theme(legend.position="bottom"), 
  ggplot() + geom_sf(data = na.omit(tractsBC19), aes(fill = incomeContext)) +
    scale_fill_manual(values = c("#25CB10", "#FA7800"), name="Income Context") +
    labs(title = "Income Context") +
    mapTheme() + theme(legend.position="bottom"))
```


```{r income_mape_table, cache = TRUE, message = FALSE, warning = FALSE}
st_join(bothRegressions, tractsBC19) %>% 
  filter(!is.na(incomeContext)) %>%
  group_by(Regression, incomeContext) %>%
  summarize(mean.MAPE = scales::percent(mean(price.APE, na.rm = T))) %>%
  st_drop_geometry() %>%
  spread(incomeContext, mean.MAPE) %>%
  kable(caption = "Test set MAPE by neighborhood income context")
```


<br />
<br />

## 7. Conclusion
[placeholder of paragraphs]







